---
title: "HW1_Summary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HW1_Summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#Dataset in the package
```{r setup}
library(bis557)
library(DT)
datatable(lm_patho)
```
Above is the `lm_patho` dataset in package bis557. This dataset is used to check the validity of linear_model() function in a tougher case.

#Functions in the package
We have 2 functions in this package: `linear_model()` which is used to fit the linear model by QR decomposition, and `my_gd()`, which is used to implement the gradient descent to find the fitted coefficients.

Below we use the dataset iris as example:
```{r}
linear_model(Sepal.Length ~ ., iris)
```
Also we can apply a tough data set `lm_patho` to this function:
```{r}
linear_model(y ~ ., lm_patho)
```
 We can see that the output coefficients is very close to the output of `lm()` function.
